Towards caring 15-minute neighbourhoods

### 2024-05-28 data/03-clustering-and-profiling.qmd
Added more variables to "HAM_census_21.rda" using 01-census-and-traveltimes.qmd. Played around in 03-clustering-and-profile.qmd

Select data for classification:
```{r}
df <- access_15_wide |>
  transmute(scluster = factor(scluster),
            LICO = `v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`)
```

>>>>

Select data for classification:
```{r}
df <- access_15_wide |>
  transmute(scluster = factor(scluster),
            LICO = `v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`,
            Unemployed = `v_CA21_6501: Unemployed`/(`v_CA21_6498: Employed` + `v_CA21_6501: Unemployed` + `v_CA21_6504: Not in the labour force`),
            Income_gini = `v_CA21_1142: Gini index on adjusted household after-tax income`,
            Vis_minority = `v_CA21_4875: Total visible minority population`/ (`v_CA21_4914: Not a visible minority`+ `v_CA21_4875: Total visible minority population`),
            Age_0to14_and_Age_65plus =( `v_CA21_11: 0 to 14 years` + `v_CA21_251: 65 years and over`) / (`v_CA21_11: 0 to 14 years` + `v_CA21_68: 15 to 64 years` + `v_CA21_251: 65 years and over`),
            One_parent_families = `v_CA21_507: Total one-parent families` / (`v_CA21_507: Total one-parent families` + `v_CA21_500: Total couple families`) )

# # examining correlations
# cor(df |> select(-c("scluster")), use = "complete.obs")
```

### 2024-05-28 paper/Manuscript.qmd
Added more variables to "HAM_census_21.rda" using 01-census-and-traveltimes.qmd. Played around in 03-clustering-and-profile.qmd

- Next: describe hamilton's land-use context; communities which are urban, rural, suburban. Also describe the residential parcels summary statistics, ~140,000 points, aggregated at the level of Canadian Census DA. See population descriptive plots in @fig-Fig1:
```{r plot-figure1, out.width=600, out.height=600}
#| label: fig-Fig1
#| fig-cap: "The number of residential parcels per DA in 2020 (top) and the rate of population per dwelling per DA (middle) and prevalence of Low-income cut-offs after-tax (LICO-AT) per DA (bottom) retrieved from the 2021 Canadian Census. All scales in quartiles. Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/data/pop_descriptive_plots.tiff"))
```

@@

```{r plot-figure2, out.width=600, out.height=600}
#| label: fig-Fig2
#| fig-cap: "The locations of care destinations in Hamilton separated by the author-generated categories of: dependent-, errand-, grocery- and health- centric care categories. Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/data/care_categories_plot.tiff"))
```

@@

# Results
<!--Initial thoughts on results:
-Initial results seem optimistic. With SOM , method will classify neighbourhoods into these clusters we then tag/name. 
- A part of this - spatial knowledge of people doesn't match the dataset. i.e. all stores are not equal, they have qualities that make them less/more accessibly. 
- But we can speak in potentials. They facilities are there and they could potentially be improved. -->
```{r plot-figure3, out.width=600, out.height=600}
#| label: fig-Fig3
#| fig-cap: "The cumulative opportunity measure. The number of care destinations that can be reached, per DA, within 15 min walk. Visualised as 'caring' (based on the opportunity reachable number; Caring is approximately the median) and if complete (if all care categories can be reached; this is an arbitrary selection). Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_copp_plot.tiff"))
```

```{r plot-figure4, out.width=600, out.height=600}
#| label: fig-Fig4
#| fig-cap: "The diversity measure based on the _specific type_ of destination cumulative opportunity DA measure. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_entropy_copp_plot.tiff"))
```

```{r plot-figure5, out.width=600, out.height=600}
#| label: fig-Fig5
#| fig-cap: "A bivariate map of the cumulative opportuninities map (aggregated per category at DA - fig3) and the diversity measure (fig 4). Amenity diverse = diviersity index above the region median. Caring/Complete = accessibility value above the region median. The Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_bivariate_entropy_copp_plot.tiff"))
```

>>>>

- Next: describe hamilton's land-use context; communities which are urban, rural, suburban. Also describe the residential parcels summary statistics, ~140,000 points, aggregated at the level of Canadian Census DA. See population descriptive plots in @fig-Fig1:
```{r plot-figure1, out.width=600, out.height=600}
#| label: fig-Fig1
#| fig-cap: "The number of residential parcels per DA in 2020 (top) and the rate of population per dwelling per DA (middle) and prevalence of Low-income cut-offs after-tax (LICO-AT) per DA (bottom) retrieved from the 2021 Canadian Census. All scales in quartiles. Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/pop_descriptive_plots.tiff"))
```

@@

```{r plot-figure2, out.width=600, out.height=600}
#| label: fig-Fig2
#| fig-cap: "The locations of care destinations in Hamilton separated by the author-generated categories of: dependent-, errand-, grocery- and health- centric care categories. Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/care_categories_plot.tiff"))
```

@@

# Results
<!--Initial thoughts on results:
-Initial results seem optimistic. With SOM , method will classify neighbourhoods into these clusters we then tag/name. 
- A part of this - spatial knowledge of people doesn't match the dataset. i.e. all stores are not equal, they have qualities that make them less/more accessibly. 
- But we can speak in potentials. They facilities are there and they could potentially be improved. -->
```{r plot-figure3, out.width=600, out.height=600}
#| label: fig-Fig3
#| fig-cap: "The cumulative opportunity measure. The number of care destinations that can be reached, per DA, within 15 min walk. Visualised as 'caring' (based on the opportunity reachable number; Caring is approximately the median) and if complete (if all care categories can be reached; this is an arbitrary selection). Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_copp_plot.tiff"))
```

```{r plot-figure4, out.width=600, out.height=600}
#| label: fig-Fig4
#| fig-cap: "The diversity measure based on the _specific type_ of destination cumulative opportunity DA measure. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_entropy_copp_plot.tiff"))
```

```{r plot-figure5, out.width=600, out.height=600}
#| label: fig-Fig5
#| fig-cap: "A bivariate map of the cumulative opportuninities map (aggregated per category at DA - fig3) and the diversity measure (fig 4). Amenity diverse = diviersity index above the region median. Caring/Complete = accessibility value above the region median. The Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_bivariate_entropy_copp_plot.tiff"))
```
